---
title: "Mental Health in Tech"
author: "Shawn D'Souza"
date: "June 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

**
# Introduction
**
Assessment of mental health is a leading factor in today's society.  It goes doubly in this day and age where has taken a higher priority now more than ever. There are so many things that one can take into factor when looking at what makes people stress in their line of work.

This survey was conducted to find what people think of their work and the various factors that they feel makes them stressed or not.   


Before I dive into the technical part, I would like to thank the following people from who I have gained valuable insights while implementing this project:



<a = href "https://www.kaggle.com/lislejoem">Lislejoem</a>
<a = href "https://www.kaggle.com/lislejoem/recognition-of-importance-of-mental-health-in-us"></a>

<a = href "https://www.kaggle.com/ekim01">Esther Kim</a>
<a = href "https://www.kaggle.com/ekim01/predictors-of-mental-health-illness"></a>

<a = href "https://www.kaggle.com/devisangeetha">DSangeetha</a>
<a = href "https://www.kaggle.com/devisangeetha/analysis-on-survey-of-mental-health"></a>

<a = href "https://www.kaggle.com/diegocalvo">Diego Calvo</a>
<a = href "https://www.kaggle.com/diegocalvo/data-mining-of-mental-health"></a>

<a = href "https://www.kaggle.com/benjaminlott">BenjaminLott</a>
<a = href "https://www.kaggle.com/benjaminlott/machine-learning-for-mental-health-7-methods"></a>



With credit given where credit is due, this Exploratory Data Analysis (EDA) encompasses with the amalgamation of concepts related to cleaning, defining and analyzing data to obtain meaningful insights. So let's dive in shall we?

# 1. Load the packages

```{r Load_Data_Packages,eval=TRUE,echo=TRUE,warning=FALSE, message=FALSE,results='hide'}

# Function: installing and loading of packages
install_load <- function (packages)  {   
   
  # Start loop to determine if each package is installed
   for(package in packages){

       # If package is installed locally, load
       if(package %in% rownames(installed.packages()))
          do.call('library', list(package))

       # If package is not installed locally, download, then load
       else {
          install.packages(package, dependencies = TRUE)
          do.call("library", list(package))
       }
   } 
}

# Generic libraries loading
libs <- c("ggplot2", "maps", "plotly", "plyr", "dplyr", "rworldmap","stringr","lubridate", "plotly", "reshape2", "magrittr", "ggthemes", "tidyr", "DT", "lubridate","RColorBrewer")
install_load(libs)

# Specific methods libraries loading
libs.methods <- c("C50", "lattice", "caret", "nnet", "e1071","Matrix", "foreach","glmnet","C50","randomForest","ipred","rpart")
install_load(libs.methods)

# Data loading
#data <- read.csv("../input/survey.csv")
data <- read.csv("survey.csv")
data
```

# Determine the different values in each of the categories
```{r categorical_splits, eval=TRUE,echo=TRUE}


data <- data[ , !(names(data) %in% "Timestamp")]


#Finding the NA Categories
sapply(data, function(x) sum(is.na(x)))


data %<>% dplyr::select(-c(comments, state, work_interfere))

categories <- names(data)
print(categories)

for (val in categories){ 
    print(unique(data[val])) 
    
}

```

#Deep diving into the Genders

This was very interesting to work with as this showed the diversity of data and how to clean data to split into isnights

```{r}
#Diving into the genders
male_str = c("male", "m", "male-ish", "maile", "mal", "male (cis)", "make", "male ", "man","msle", "mail", "malr","cis man", "Cis Male", "cis male")
trans_str = c("trans-female", "something kinda male?", "queer/she/they", "non-binary","nah", "all", "enby", "fluid", "genderqueer", "androgyne", "agender", "male leaning androgynous", "guy (-ish) ^_^", "trans woman", "neuter", "female (trans)", "queer", "ostensibly male, unsure what that really means")           
female_str = c("cis female", "f", "female", "woman",  "femake", "female ","cis-female/femme", "female (cis)", "femail")

data$Gender %<>% str_to_lower()
data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% trans_str) "trans" else x )
data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% male_str) "male" else x )
data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% female_str) "female" else x )

```





```{r}
print(unique(data$Gender))
data %<>% filter(Gender != "a little about you")
data %<>% filter(Gender != "p")
print(unique(data$Gender))
```


# Clean the Age category: Since there are anomalies within the age bracket

```{r}
# Age categorization 
data$Age <-
  ifelse(data$Age < 18, "",   
  ifelse(data$Age > 100, "", data$Age)) %>% 
  as.numeric()

```
```{r}
#TechData <- data %>%
  #filter(tech_company == "Yes" )& # Making sure dataset includes only those in a tech company
         #format(Timestamp, "%Y") == 2014 & # Making sure dataset includes only responses from 2014
         
summary(data)

```


```{r}

# Age categorization
#Finding the NA Categories
sapply(data, function(x) sum(is.na(x)))
data <- data[!is.na(data$Age),]
data <- data[!is.na(data$self_employed),]
```

```{r}
print(head(data))
```
# Basic Exploratory Data Analysis


```{r}
#Based on people taking treatments for their stress
for(i in 1:length(data)){
  aux <- prop.table(table(data$treatment, data[,i]), 1)*100 
  percent <- round(max(abs(aux[1,]-aux[2,])), digits = 2)

  if(percent > 10 & percent < 99){
    
    # Data preparing to visualization
    aux <- prop.table(table(data$treatment, data[,i]), 1)*100 
    nom <- colnames(aux)
    taken_treatment <- c(rep("Yes",ncol(aux)),rep("No",ncol(aux)))
    val <- append(aux[1,], aux[2,])
    data.aux<-data.frame(nom=nom,type=taken_treatment ,val=val)
    
    # Use of the library ggplot2 to data visualization 
    g <- ggplot() + geom_bar(data=data.aux,aes(x=nom, y=val,fill=taken_treatment),stat='identity',position='dodge')+
      coord_flip() +
      labs(
        x = "Importance",
        y = "",
        title = paste("Mental Health comparation about ", names(data[i]), sep=""),
        subtitle = paste("The most different is ", percent, "%", sep=""),
        caption = "\nDetermined by matrix of covariances"
        ) %>% suppressWarnings()
    print(g)
  }

}
```
```{r}
data1=data
# Age categorization#
data1$Age<-cut(data1$Age, breaks = c(16, 33, 47, 100), labels = c('Millennials', 'Gen X', 'Baby Boomers'))

age_group <-data1%>%
  group_by(Age) %>%
  dplyr::summarize(count = n())
# Visualize the number of subjects in each Age Group  
ggplot(age_group, aes(x = Age, y = count, fill = Age)) +  
  geom_bar(stat = "identity", alpha = 0.5) +
  xlab("Age Group") + 
  ylab("No of People") + 
  ggtitle("Comparing Age Group in the 2014 Mental Health in Tech Survey")
```




```{r}
country <- data1$Country
ill <- data1$treatment
data1.aux <- data.frame(country,ill)

# Data frame definition
data1.result <- ddply(data1.aux,.(country,ill), nrow)

# Frecuency accumulation of treated people
frec <- data.frame(id = data1.result$country
                   , value = data1.result$V1)

frec <- mutate(group_by(frec,id), cumsum=cumsum(value))

data1.all <- data.frame(id=frec$id,num=frec$cumsum)
data1.all <- data1.all[with(data1.all, order(-data1.all$num)), ]
data1.all <- data1.all[!duplicated(data1.all$id),]
data1.all <- data1.all[with(data1.all, order(data1.all$id)), ]

# Putting treated people in negative way
for (i in 1:length(data1.result$country)){
    if(data1.result$ill[i] =="Yes"){
      data1.result$V1[i] <- data1.result$V1[i]*-1
    }
}

# Frecuency accumulation of treated people
frec <- data.frame(id = data1.result$country
                          , value = data1.result$V1)

frec <- mutate(group_by(frec,id), cumsum=cumsum(value))

data1.aux <- data.frame(id=frec$id,num=frec$cumsum)
data1.aux <- data1.aux[with(data1.aux, order(-data1.aux$num)), ]
data1.aux <- data1.aux[!duplicated(data1.aux$id),]
data1.aux <- data1.aux[with(data1.aux, order(data1.aux$id)), ]

# Making relative treated people
data1 <- data.frame(id=data1.all$id)
data1$id <- data1.aux$id
data1$num <- (data1.aux$num / data1.all$num )
data1$category <- data1$num

# Categorization of treated people by geographic location
data1$category  <- cut(data1$num, breaks=c(-Inf, -0.75, -0.5, -0.35, 0, 0.35, 0.5, 0.75, Inf))
levels(data1$category) <- c("-70%","-50%","-30%","-15%","+15%","+30%","+50%","+70%")

d <- data.frame( country=data1$id, value=data1$cate)
n <- joinCountryData2Map(d, joinCode="NAME", nameJoinColumn="country")

## 39 codes from your data successfully matched countries in the map
## 0 codes from your data failed to match with a country code in the map
## 204 codes from the map weren't represented in your data

# Visualization of treated people by geographic location
mapCountryData(n, nameColumnToPlot="value", mapTitle="World" ,catMethod="categorical",
               colourPalette=c('red','blue'),missingCountryCol="grey", aspect =0)
```

Athough this is a brief scraping, this shows the bare-bones of data analysis with R. 

All in all R is still a nuanced concept that I would love to explore fruther into. 



# Do Reach Out!

If you found something that catches your eye or just have a chat about data science topics in general, I will be more than happy to connect with you on:

LinkedIn: https://www.linkedin.com/in/shawn-dsouza/

My Website: https://shawndsouza29.wixsite.com/portfolio

This notebook will always be a work in progress. Please leave any comments about further improvements to the notebook! Any feedback or constructive criticism is greatly appreciated. Thank you guys!



